Measurement of deeply virtual Compton scattering off Helium-4 with CLAS at Jefferson Lab

### 2020-03-25 eg6-letters/long-paper/prc-dvcshelium.tex
Add corrections to long paper

In the past few decades, the study of the proton structure has made significant progress
thanks to the theoretical and experimental developpements with three dimensional 
structure functions \cite{Anselmino:2015uka}. These studies, which have focused on generalized
parton distributions (GPD) and transverse momentum dependent parton distribution functions (PDF)
can be generalized to the nucleus and offer a unique opportunity to revisit the quark structure 
of the nucleus with an original perspective~\cite{Dupre:2015jha}. This new approach is 
particularly needed as the quark structure of the nucleus remains today the subject of numerous
controvercies. Indeed, while much progress has been made in measuring the nuclear
PDFs, their shape remains fundamentaly unexplained \cite{Norton:2003cb,Malace:2014uea,Hen:2016kwk}.

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin assymetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions in the nuclear 
medium and offers a unique opportunity to make a measurement free of final state interactions 
(FSI). Moreover, with a nuclear targets of spin-0 the extraction of the GPD from the DVCS data 
is significantly simplified with the reduction to a single GPD intervening in the process at leading order. The 
measurement of the nuclear DVCS is however challenging experimentally, first attempts from 
the HERMES collaboration \cite{Airapetian:2009cga} lead to controversial conclusions. We present here
in detail the more recent measurements from the CLAS collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the nuclear DVCS CLAS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

Experimentally, another process is indistinguishable from the DVCS, the Bethe-Heitler (BH) 
process in which % TODO add BH figure ?
the final state photon is emitted by the scattering lepton rather than the hadron. In this case, the 
photon-hadron interaction is the same as in elastic scattering and depends on the
target FFs rather than its GPDs. The DVCS and BH processes are experimentally indistinguishable as they
have identical final states, such that they interfere in the squared amplitude of the
exclusive photo-production process:
\begin{equation}
	|T|^2 = |T_{DVCS}|^2 + |T_{BH}|^2 + T_{DVCS}^* T_{BH} + T_{DVCS}T_{BH}^*. 
\end{equation}
The interference terms increase significantly the cross section in specific parts of the phase space
and lead to significant beam spin asymmetries (BSA), which are the focus of the measurements presented 
here.

@@

\begin{figure}[tbp!]
\center
\includegraphics[width=9cm]{fig1/finalplane.png}
	\caption{Illustration of the scattering (or leptonic) and production (or hadronic) planes
	in the DVCS process.} %TODO upgrade to a figure with phi_S
\label{fig:PhiAngle}
\end{figure}

@@

The first reaction measured in the experiment is the coherent electro-production of a photon on helium
$e+^4\!\!He \rightarrow e+\gamma+^4\!\!He$ at high 4-momentum transfer squared ($Q^2$). The 
leading order diagram of the nuclear coherent DVCS is represented in Fig. \ref{fig:CohDiag}.
In our experiment, we focused on the measurement of the beam-spin asymmetry (BSA) noted $A_{LU}$ with 
$L$ for the longitudinally-polarized electron beam and $U$ the unpolarized target and defined as:  
\begin{equation}
A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
              {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
  \label{eq:BSA}
\end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the differential cross section for a positive 
(negative) beam helicity. At leading order and leading twist, the BSA can be expressed as \cite{Kirchner:2003wt}:        
\begin{eqnarray}
A_{LU}& =& \frac{x_A(1+\epsilon^2)^2}{y} \, s_1^{INT} \sin(\phi) \, 
\bigg/ \, \bigg[ \, \sum_{n=0}^{n=2}c_n^{BH}\cos{(n\phi)} +  \\
& & \frac{x_A^2 t {(1+\epsilon^2)}^2}{Q^2} P_1(\phi) P_2(\phi) \, c_0^{DVCS} + 
\frac{x_A (1+\epsilon^2)^2}{y} \sum_{n=0}^{n=1} c_n^{INT} \cos{(n\phi)} \bigg].  \nonumber 
\label{eq:coh_BSA}
\end{eqnarray}
where $\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ are the BH 
propagators. The factors: $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and 
$s_1^{INT}$ are the Fourier coefficients of the BH, the DVCS and the 
interference amplitudes for a spin-zero target. The explicit 
expressions of these coefficients, which have been derived based on the work of 
Kirchner and Müller \cite{Kirchner:2003wt}, can be found in Appendix \ref{app:Helium_cross_section}.

@@

This formula can be expressed in a simplified manner for a spin-0 target as \cite{Belitsky:2008bz}:
\begin{equation}
A_{LU}(\phi) = \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the CFF $\mathcal{H}_{A}$ associated to the GPD $H_A$ of the 
spin-0 nucleus. The 
$\alpha_{i}$ factors are $\phi$-dependent kinematical terms that depend on the 
nuclear form factor $F_A$ and the independent variables $Q^2$, $x$ and $t$.  
These factors have the following simplified expressions:
%\small
\begin{eqnarray}
	\label{eq:alpha1}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
	\label{eq:alpha4}
\end{eqnarray}
%\normalsize
where $S_{++}(1)$, $C_{++}(0)$, and $C_{++}(1)$ are the Fourier harmonics in the 
leptonic tensor \cite{Belitsky:2008bz}, their explicit expression can be found in 
appendix \ref{sec:eq}, and $x_{A} = \frac{M_{p}\cdot x}{M_{^4\!He}}$.

@@

The incoherent nuclear DVCS process, is the DVCS off a bound nucleon in a nucleus
as represented in Fig. \ref{fig:InCohDiag} for an helium-4 target. The remnants of 
the nucleus ($X$) contain only the missing three nucleons. 
The theory for incoherent DVCS on the nucleon is largely based on the free proton theory
already reviewed many times in the litterature~\cite{Diehl:2003ny,Belitsky:2005qn,Guidal:2013rya}. 
Two important differences need to be accounted for however, the different initial state and the addition of 
final state effects. In the initial state, the intrinsic Fermi motion of the nucleons in the nucleus 
lead to an uncertainty on the exact kinematic of the reaction. Moreover, in general the nucleon 
is in an off-shell state that is not exactly identical to its final state. In the final state, 
interactions between the outgoing nucleon from the DVCS reaction and remnants of the nuclear 
target are possible. The latter leading to contamination from other channels, in particular 
charge exchange processes can lead to a large contribution from such background reaction.

@@

\begin{figure}[tbp!]
\center
	\includegraphics[width=10.5cm]{fig3/HERMES_BSA.png}
	\caption{The sinus moment of the BSA as a function of $-t$ measured by HERMES
	on a series of nuclei \cite{Airapetian:2009cga}.}
\label{fig:HERMES1}
\end{figure}

@@

The CLAS experiment had for main objective to explore the coherent DVCS on helium-4 and assess if
the predicted BSA increase could be observed and if GPDs could be extracted from it. In order to
perform this measurement however, several instrumentation challenges need to be passed. First, to
measure the scattered electron and the low angle photon of the DVCS, we used CLAS in its 
DVCS setup, $i.e.$ with the addition of a low angle calorimeter and a solenoid magnet. Second, the
helium recoil need to be measured to ensure the exclusivity in the coherent channel. In this section, we 
will rapidly describe the particular detector setup used in our experiment.

\subsection{The CLAS detector}

The CLAS spectrometer \cite{Mecking:2003zu} is installed in the hall-B of the JLab main 
accelerator. It has for main objective to study the multi-particles final states
that cannot be reconstructed with arm spectrometers. Its historic use to measure DVCS in many different 
configurations made it an ideal place for this new DVCS measurement.  
CLAS is composed of six identical radial 
sectors separated by a toroidal magnet and each sector is made of four detectors 
as shown in Fig. \ref{fig:CLAS}.  Three regions of drift chambers \cite{Mestayer:2000we} 
are placed in the solenoid area to reconstruct the charge particle's tracks and 
calculate their momentum. Then, an array of scintillators is placed to measure 
a precise time for each track and perform time-of-flight measurement 
\cite{Smith:1999ii}. These detectors cover the polar angle from 8 to 142 degrees. 
This is completed by two important forward detectors, from 8 
to 45 degrees for electron identification, a Cerenkov counter 
\cite{Adams:2001kk} and an electromagnetic calorimeter \cite{Amarian:2001zs}.

@@

The inner calorimeter, illustrated in Fig. \ref{fig:IC} is a homogeneous 
calorimeter composed 424 lead tungstate 
(PbWO) crystals read out by $5 \times 5$ mm$^2$ avalanche photo-diodes (APDs). 
It covers angles from 4 to 15 degrees. However, placing a detector at such low angles makes it 
particularly sensitive to the Moller electrons scattered from the target at small angles.

@@

To protect the calorimeter from the low energy background a 5~T solenoid is 
placed around the target to form
a magnetic shield. Thanks to this field, low energy charged particles (particularly 
electrons) curl around the beam line 
and never make it to the calorimeter or other CLAS detectors as illustrated 
by simulation results presented 
in Fig. \ref{fig:Solenoid}. This allows to run much higher luminosity experiments,
a necessity for low rates processes like DVCS.

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV/c at our beam energy. Therefore, a specific detector was 
needed to detect them. To design the present setup, large inspiration was drawn from the 
Bonus setup that also used a GEM based RTPC \cite{Fenker:2008zz}, installed in CLAS to 
detect slow protons out of a deuterium target \cite{Baillie:2011za}. In such an RTPC the 
charges are projected toward large radii 
rather than towards the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere \cite{Dupre:2017upj}, here is a summary of key elements.

In order to detect the recoil helium nuclei from a DVCS reaction, we first need 
to ensure that it will come out of the target. For this, we used a light straw 
target made of a thin kapton wall of 27 $\mu$m filled with up to 6 atm of 
helium. The entrance and exit windows are thin aluminum foils and an helium 
bag is placed downstream of the target to avoid interaction with air in the gap 
between the target and the beamline vacuum. The cylindrical 
chamber surrounds the target as illustrated in Fig. \ref{fig:RTPCGlobal}, we 
list the elements composing it based on their radii:
\begin{itemize}
	\item Up to a radius of 3 mm the pressurized helium target.
	\item From 3 to 20 mm a keep-out zone filled with 1 atm of helium to 
		minimize the production of secondaries. 
	\item At 20 mm a grounded foil made of 4 $\mu$m aluminized Mylar to 
		isolate the chamber from the beam line 
		region and collect charges. It also serves to separate gas regions.
	\item From 20 to 30 mm a dead zone to separate the ground from the cathode
		filled with the drift gas, a mix of neon and dimethyl ether (DME) in 
		80/20 proportions.
	\item At 30 mm the cathode foil made of 4 $\mu$m aluminized Mylar.
	\item From 30 to 60 mm the drift region filled with the drift gas.
	\item From 60 to 69 mm the amplification regions, filled with drift gas, 
		with GEM foils placed at 60, 63 and 66 mm.
	\item At 69 mm the collection pads connected to the preamplifers placed 
		directly outside the chamber.
\end{itemize}

@@

The calibration of the detector has been performed with a 
dependence on $z$, the position along the beam line axis, due to 
variations in the magnetic fields. To perform this calibration we took  
dedicated data at 1.2 GeV beam energy. In 
this data set, we were able to select elastic events, for which the kinematic 
of the helium recoil can be calculated from the electron kinematic and 
directly compared to the measurement in the RTPC. This comparison helped to 
map the correspondence between time and position in the chamber and determine 
the drift path of electrons. A more detailed description of the calibration process 
is available in \cite{Dupre:2017upj}.

@@

The electrons are detected with the baseline CLAS detectors, the drift chamber measures the kinematic 
of the electron and the large signal in both the Cerenkov counter and electromagnetic calorimeter
provide the identification. A signal of good quality is also required in the 
time-of-flight system, which serves as a time reference for other detectors. In particular, it 
serves for the identification of the protons. Several 
fiducial cuts are applied to ensure that particles did not go through part of the inner calorimeter 
or the solenoid, as well as to reject the edges of the detectors, where efficiency is rapidly 
decreasing. Kinematic corrections are also applied to the electrons and protons to correct for energy loss
and biases in calibration, they are at subpercent level except for protons below 500 MeV/c for which they
go up to 10\% at the detection limit of 200 MeV/c.

@@

Finally, we select events that contain a single electron, a high energy photon ($E>2$ GeV) and
either a helium or a proton. We perform a vertex selection cut on the two charged particles 
to ensure they originate from the same vertex, inside the target, thus rejecting target windows.
Moreover, since we are aiming to study deep processes occurring at the partonic level, we select
$Q^2>1$ GeV$^2$. Also, the transferred momentum squared to the recoil $^{4}He$ is bound by
   a minimum value based on basic energy momentum conservation:
\begin{equation}
   t_{min} = - Q^{2} \frac{2(1- x_{A})(1 - \sqrt{1 + \epsilon ^{2}}) + \epsilon 
   ^{2}}{4 x_{A}(1-x_{A})+ \epsilon ^{2}},\\
\end{equation}
where $\epsilon ^{2} = \frac{4M^{2}_{^4\!He}x^{2}_{A}}{Q^{2}}$. For incoherent DVCS, 
$x_A$ is replaced by $x$ and $M_{^4\!He}$ by $M_{p}$.

@@

To select exclusive events and suppress the backgrounds, we define several 
exclusivity cuts based on energy and momentum conservation. In principle, 
a selection based on two or three variables can be used to guaranty the 
exclusivity of the process. However, in such experiment, where particles 
are detected at very different energies and with very different detector 
precision, we chose to over constrain the selection by using seven 
variables. The seven variables are defined as follow for the coherent 
DVCS (replace helium by proton for the incoherent case):
\begin{itemize}
	\item Co-planarity ($\Delta \phi$) of the virtual photon, the real photon and
		the recoil helium, defined as the angle between the helium, virtual photon
		plane and the virtual photon, real photon plane;
	\item Missing energy of the full system;
	\item Missing mass of the full system;
	\item Missing transverse momentum of the full system;
	\item Missing mass of the system electron-helium, ignoring the photon;
	\item Missing mass of the system electron-photon, ignoring the recoil helium;
	\item Co-linearity ($\theta$) of the measured photon with the missing momentum of the 
		electron-helium system.
\end{itemize}

In the analysis, we apply selection cuts based on a fit of the exclusive peak at 3$\sigma$ around 
the mean value for each variable. This automatic method helps to avoid any bias 
in the selection of the events. The selection of coherent DVCS with these variables is illustrated in 
Fig. \ref{fig:CohExcCuts}. We note on these distributions only few minor anomalies, where the 
distribution have some asymmetries. These are linked with the detector resolutions, that impact
differently the kinematic variables. The selection of incoherent DVCS is presented in 
Fig. \ref{fig:IncExcCuts}, with two main differences: larger distributions overall and more
offset distributions. The former is largely due to Fermi motion, but simulations have shown
that this effect is not strong enough to reproduce these distribution widths and final 
state interactions must play a role as well. The later is caused by slight detector 
misalignment between CLAS sectors and are within the levels obtained with free proton 
targets to which they can be directly compared (see \cite{HirlingerSaylor:2018bnu} for instance).

\begin{figure}[tbp!]
\center
\includegraphics[trim=20 10 20 5,clip,width=15cm]{fig3/all_coh_exc_cuts.pdf}
	\caption{Overlay of the events distributions before (black) and after (blue line filled
	in Grey) the exclusivity cuts used to select coherent DVCS represented by the 
	red dashed lines. The histograms are for the seven variables described in the text, 
	plus the missing $P_x$ and $P_y$, in order left to right and top to bottom.}
\label{fig:CohExcCuts}
\end{figure}

>>>>

In the past few decades, the study of the proton structure has made significant progress
thanks to the theoretical and experimental developments with three dimensional 
structure functions \cite{Anselmino:2015uka}. These studies, which have focused on generalized
parton distributions (GPD) and transverse momentum dependent parton distribution functions (PDF)
can be generalized to the nucleus and offer a unique opportunity to revisit the quark structure 
of the nucleus with an original perspective~\cite{Dupre:2015jha}. This new approach is 
particularly needed as the quark structure of the nucleus remains today the subject of numerous
controversies. Indeed, while much progress has been made in measuring the nuclear
PDFs, their shape remains fundamentally unexplained \cite{Norton:2003cb,Malace:2014uea,Hen:2016kwk}.

In nuclei, the GPDs can be probed conveniently through the measurement of the
spin asymmetries generated by the deeply virtual Compton scattering (DVCS) 
process~\cite{Diehl:2003ny,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}. The measurement of the exclusive
production of a photon limits the possibilities of final state interactions in the nuclear 
medium and offers a unique opportunity to make a measurement free of final state interactions 
(FSI). Moreover, with a nuclear targets of spin-0 the extraction of the GPD from the DVCS data 
is significantly simplified with the reduction to a single GPD intervening in the process at leading order. The 
measurement of the nuclear DVCS is however challenging experimentally, first attempts from 
the HERMES collaboration \cite{Airapetian:2009cga} lead to controversial conclusions. We present here
in detail the more recent measurements from the CLAS collaboration, which has been already partially presented
in two short letters \cite{Hattawy:2017woc,Hattawy:2018liu}. We extend in this article 
the description of the nuclear DVCS CLAS experiment, detail the methods used for the
data analysis and produce the complete experimental results for each channel measured.

@@

Experimentally, another process is indistinguishable from the DVCS, the Bethe-Heitler (BH) 
process in which 
the final state photon is emitted by the scattering lepton rather than the hadron. In this case, the 
photon-hadron interaction is the same as in elastic scattering and depends on the
target FFs rather than its GPDs. The DVCS and BH processes are experimentally indistinguishable as they
have identical final states, such that they interfere in the squared amplitude of the
exclusive photo-production process:
\begin{equation}
	|T|^2 = |T_{DVCS}|^2 + |T_{BH}|^2 + T_{DVCS}^* T_{BH} + T_{DVCS}T_{BH}^*. 
\end{equation}
The interference terms increase significantly the cross section in specific parts of the phase space
and lead to significant beam spin asymmetries (BSA), which are the focus of the measurements presented 
here.

@@

\begin{figure}[tbp!]
\center
\includegraphics[width=9cm]{fig1/finalplane.png}
	\caption{Illustration of the scattering (or leptonic) and production (or hadronic) planes
	in the DVCS process.} 
\label{fig:PhiAngle}
\end{figure}

@@

The first reaction measured in the experiment is the coherent electro-production of a photon on helium
$e+^4\!\!He \rightarrow e+\gamma+^4\!\!He$ at high 4-momentum transfer squared ($Q^2$). The 
leading order diagram of the nuclear coherent DVCS is represented in Fig. \ref{fig:CohDiag}.
In our experiment, we focused on the measurement of the beam-spin asymmetry (BSA) noted $A_{LU}$ with 
$L$ for the longitudinally-polarized electron beam and $U$ the unpolarized target and defined as:  
\begin{equation}
A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
              {d^{5}\sigma^{+} + d^{5}\sigma^{-}},
  \label{eq:BSA}
\end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the differential cross section for a positive 
(negative) beam helicity. At leading order and leading twist, the BSA can be expressed as \cite{Kirchner:2003wt}:        
\begin{eqnarray}
\label{eq:coh_BSA}
A_{LU}& =& \frac{x_A(1+\epsilon^2)^2}{y} \, s_1^{INT} \sin(\phi) \, 
\bigg/ \, \bigg[ \, \sum_{n=0}^{n=2}c_n^{BH}\cos{(n\phi)} +  \\
& & \frac{x_A^2 t {(1+\epsilon^2)}^2}{Q^2} P_1(\phi) P_2(\phi) \, c_0^{DVCS} + 
\frac{x_A (1+\epsilon^2)^2}{y} \sum_{n=0}^{n=1} c_n^{INT} \cos{(n\phi)} \bigg].  \nonumber 
\end{eqnarray}
where $\mathcal{P}_1(\phi)$ and $\mathcal {P}_2(\phi)$ are the BH 
propagators. The factors: $c_{0,1,2}^{BH}$, $c_0^{DVCS}$, $c_{0,1}^{INT}$ and 
$s_1^{INT}$ are the Fourier coefficients of the BH, the DVCS and the 
interference amplitudes for a spin-zero target. The explicit 
expressions of these coefficients, which have been derived based on the work of 
Kirchner and Müller \cite{Kirchner:2003wt}, can be found in Appendix \ref{app:Helium_cross_section}.

@@

This formula can be expressed in a simplified manner for a spin-0 target as \cite{Belitsky:2008bz}:
\begin{equation}
A_{LU}(\phi) = \frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( 
\Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}
\label{eq:A_LU-coh}
\end{equation}
where $\Im m(\mathcal{H}_{A})$ and $\Re e(\mathcal{H}_{A})$ are the imaginary 
and real parts of the CFF $\mathcal{H}_{A}$ associated to the GPD $H_A$ of the 
spin-0 nucleus. The 
$\alpha_{i}$ factors are $\phi$-dependent kinematical terms that depend on the 
nuclear form factor $F_A$ and the independent variables $Q^2$, $x$ and $t$.  
These factors have the following simplified expressions:
\begin{eqnarray}
	\label{eq:alpha1}
   \alpha_0 (\phi) & = &\frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi)\\
    \alpha_1 (\phi) & = & c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi)\\ 
   \alpha_2 (\phi) & = & \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right)\\
\alpha_3 (\phi) &=& \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
	\label{eq:alpha4}
\end{eqnarray}
where $S_{++}(1)$, $C_{++}(0)$, and $C_{++}(1)$ are the Fourier harmonics in the 
leptonic tensor \cite{Belitsky:2008bz}, their explicit expression can be found in 
appendix \ref{sec:eq}, and $x_{A} = \frac{M_{p}\cdot x}{M_{^4\!He}}$.

@@

The incoherent nuclear DVCS process, is the DVCS off a bound nucleon in a nucleus
as represented in Fig. \ref{fig:InCohDiag} for an helium-4 target. The remnants of 
the nucleus ($X$) contain only the missing three nucleons. 
The theory for incoherent DVCS on the nucleon is largely based on the free proton theory
already reviewed many times in the literature~\cite{Diehl:2003ny,Belitsky:2005qn,Guidal:2013rya}. 
Two important differences need to be accounted for however, the different initial state and the addition of 
final state effects. In the initial state, the intrinsic Fermi motion of the nucleons in the nucleus 
lead to an uncertainty on the exact kinematic of the reaction. Moreover, in general the nucleon 
is in an off-shell state that is not exactly identical to its final state. In the final state, 
interactions between the outgoing nucleon from the DVCS reaction and remnants of the nuclear 
target are possible. The latter leading to contamination from other channels, in particular 
charge exchange processes can lead to a large contribution from such background reaction.

@@

\begin{figure}[tbp!]
\center
	\includegraphics[width=10.5cm]{fig3/HERMES_BSA.png}
	\caption{The sinus moment of the BSA as a function of $-t$ measured by HERMES
	for a series of nuclei \cite{Airapetian:2009cga}.}
\label{fig:HERMES1}
\end{figure}

@@

The CLAS experiment had for main objective to explore the coherent DVCS on helium-4 and assess if
the predicted BSA increase could be observed and if GPDs could be extracted from it. In order to
perform this measurement however, several instrumentation challenges needed to be resolved. First, to
measure the scattered electron and the low angle photon of the DVCS, we used CLAS in its 
DVCS setup, $i.e.$ with the addition of a low angle calorimeter and a 5-T solenoid magnet. Second, a
radial time projection chamber (RTPC) was installed to measure the
helium recoils and thus ensure the exclusivity of the process in the coherent channel. In this section, we 
will review the important elements of this detection setup.

\subsection{The CEBAF Large Acceptance Spectrometer (CLAS)}

The CLAS \cite{Mecking:2003zu} is installed in the hall-B of the JLab continuous electron beam 
accelerator facility (CEBAF). CLAS is specifically designed to study the multi-particles final states
that cannot be observed conveniently with arm spectrometers. Its historic use to measure DVCS in many different 
configurations made it an ideal place for this new DVCS measurement. CLAS is composed of six identical radial 
sectors separated by toroidal magnets, with each sector made of four detectors 
as shown in Fig. \ref{fig:CLAS}.  Three regions of drift chambers \cite{Mestayer:2000we} 
are placed in the solenoid area to reconstruct the charge particle's tracks and 
calculate their momentum. Then, an array of scintillators is placed to measure 
a precise time for each track and perform time-of-flight measurement 
\cite{Smith:1999ii}. These detectors cover the polar angle from 8 to 142 degrees. 
They are completed by two important forward detectors, covering the polar angle from 8 
to 45 degrees, a Cerenkov counter 
\cite{Adams:2001kk} and an electromagnetic calorimeter \cite{Amarian:2001zs}.

@@

The inner calorimeter, illustrated in Fig. \ref{fig:IC} is a homogeneous 
calorimeter composed 424 lead tungstate 
(PbWO) crystals read out by $5 \times 5$ mm$^2$ avalanche photo-diodes (APDs). 
It covers angles from 4 to 15 degrees. However, placing a detector at such low angles makes it 
particularly sensitive to the low energy Moller electrons scattered from the target.
To protect the calorimeter from this background, a 5~T solenoid is 
placed around the target to form
a magnetic shield. Thanks to this field, low energy charged particles (particularly 
electrons) curl around the beam line 
and never make it to the calorimeter or other CLAS detectors as illustrated 
by simulation results presented 
in Fig. \ref{fig:Solenoid}. This allows to run much higher luminosity experiments,
a necessity for low rates processes like DVCS.

@@

@@

The recoil helium nuclei from coherent DVCS are mostly emitted between 150 
and 200 MeV/c at our beam energy. Therefore, a specific detector was 
needed to detect them. To design the present setup, large inspiration was drawn from the 
Bonus setup that also used a GEM based RTPC~\cite{Fenker:2008zz} in CLAS to 
detect slow protons out of a deuterium target~\cite{Baillie:2011za}. In such an RTPC the 
charges are projected toward large radii 
rather than towards the endcaps, as is more traditional in time 
projection chambers. This design allows to reduce significantly the drift time and reduce the 
amount of pile-up from accidental events. The RTPC design, its operation, calibration and 
the track reconstruction have been described in more 
details elsewhere~\cite{Dupre:2017upj}, here is a summary of key elements.

In order to detect the recoil helium nuclei from a DVCS reaction, we first need 
to ensure that it will come out of the target. For this, we used a light straw 
target made of a thin kapton wall of 27 $\mu$m filled with helium at 6 atm 
pressure. The entrance and exit windows are thin aluminum foils and an helium 
bag is placed downstream of the target to avoid interaction with air in the gap 
between the target and the beamline vacuum. The cylindrical 
chamber surrounds the target as illustrated in Fig.~\ref{fig:RTPCGlobal}, we 
list the elements composing it based on their radii:
\begin{itemize}
	\item Up to a radius of 3 mm the pressurized helium target.
	\item From 3 to 20 mm a keep-out zone filled with 1 atm of helium to 
		minimize the production of secondaries. 
	\item At 20 mm a grounded foil made of 4 $\mu$m aluminized Mylar to 
		isolate the chamber from the beam line 
		region and collect charges. It also serves to separate gas regions.
	\item From 20 to 30 mm a dead zone to separate the ground from the cathode
		filled with the drift gas, a mix of neon and dimethyl ether (DME) in 
		80/20 proportions.
	\item At 30 mm the cathode foil made of 4 $\mu$m aluminized Mylar.
	\item From 30 to 60 mm the drift region filled with the drift gas.
	\item From 60 to 69 mm the amplification regions, filled with drift gas, 
		with GEM foils placed at 60, 63 and 66 mm.
	\item At 69 mm the collection pads connected to the preamplifers placed 
		directly outside the chamber.
\end{itemize}

@@

The calibration of the detector has been performed with a 
dependence on $z$, the position along the beam line axis, due to 
variations in the magnetic fields. To perform this calibration we took  
dedicated data at 1.2 GeV beam energy. In 
this data set, we were able to select elastic events, for which the kinematic 
of the helium recoil can be calculated from the electron kinematic and 
directly compared to the measurement in the RTPC. This comparison helped to 
map the correspondence between time and position in the chamber and determine 
the drift path of electrons. A more detailed description of the calibration process 
is available in~\cite{Dupre:2017upj}.

@@

The electrons are detected with the baseline CLAS detectors, the drift chamber measures the kinematic 
of the electron and the signal measured in both the Cerenkov counter and electromagnetic calorimeter
provide the identification. A signal of good quality is also required in the 
time-of-flight system, which serves as a time reference for other detectors. In particular, it 
serves for the identification of the protons. Several 
fiducial cuts are applied to ensure that particles did not go through part of the inner calorimeter 
or the solenoid, as well as to reject the edges of the detectors, where efficiency is rapidly 
decreasing. Kinematic corrections are also applied to the electrons and protons to correct for energy loss
and biases in calibration, they are at subpercent level except for protons below 500 MeV/c for which they
go up to 10\% at the detection limit of 200 MeV/c.

@@

Finally, we select events that contain a single electron, a high energy photon ($E>2$ GeV) and
either a helium or a proton. We apply a selection cut on the two charged particles 
to ensure they originate from the same vertex inside the target, thus rejecting accidentals and
events from the target windows.
Moreover, since we are aiming to study deep processes occurring at the partonic level, we select
$Q^2>1$ GeV$^2$. Also, the transferred momentum squared to the recoil $^{4}He$ is bound by
   a minimum value based on basic energy momentum conservation:
\begin{equation}
   t_{min} = - Q^{2} \frac{2(1- x_{A})(1 - \sqrt{1 + \epsilon ^{2}}) + \epsilon 
   ^{2}}{4 x_{A}(1-x_{A})+ \epsilon ^{2}},\\
\end{equation}
where $\epsilon ^{2} = \frac{4M^{2}_{^4\!He}x^{2}_{A}}{Q^{2}}$. For incoherent DVCS, 
we use a similar cut where $x_A$ is replaced by $x$ and $M_{^4\!He}$ by $M_{p}$.

@@

To select exclusive events and suppress the backgrounds, we define several 
exclusivity cuts based on energy and momentum conservation. In principle, 
a selection based on two or three variables can be used to guaranty the 
exclusivity of the process. However, in such experiment, where particles 
are detected at very different energies and with very different detector 
resolutions, we chose to over constrain the selection by using seven 
variables. The seven variables are defined as follow for the coherent 
DVCS (replace helium by proton for the incoherent case):
\begin{itemize}
	\item Co-planarity ($\Delta \phi$) of the virtual photon, the real photon and
		the recoil helium, defined as the angle between the helium, virtual photon
		plane and the virtual photon, real photon plane;
	\item Missing energy of the full system;
	\item Missing mass of the full system;
	\item Missing transverse momentum of the full system;
	\item Missing mass of the system electron-helium, ignoring the photon;
	\item Missing mass of the system electron-photon, ignoring the recoil helium;
	\item Co-linearity ($\theta$) of the measured photon with the missing momentum of the 
		electron-helium system.
\end{itemize}

In the analysis, we apply selection cuts based on a fit of the exclusive peak at 3$\sigma$ around 
the mean value for each variable. This systematic method helps to avoid any bias 
in the selection of the events. The selection of coherent DVCS with these variables is illustrated in 
Fig. \ref{fig:CohExcCuts}. We note on these distributions only few minor anomalies, where the 
distribution have some asymmetries. These are linked with the detector resolutions, that impact
some of the kinematic variables non linearly. The selection of incoherent DVCS is presented in 
Fig. \ref{fig:IncExcCuts}, with two main differences: larger distributions overall and more
offset distributions. The former is largely due to Fermi motion, but simulations have shown
that this effect is not strong enough to reproduce these distribution widths and final 
state interactions must play a role as well. The later is caused by slight detector 
misalignment between CLAS sectors and are within the levels obtained with free proton 
targets~\cite{HirlingerSaylor:2018bnu} to which they can be directly compared.

\begin{figure}[tbp!]
\center
\includegraphics[trim=20 10 20 5,clip,width=15cm]{fig3/all_coh_exc_cuts.pdf}
	\caption{Distributions of the coherent photon production events before 
	(black) and after (blue line filled
	in Grey) the exclusivity cuts used to select coherent DVCS represented by the 
	red dashed lines. The histograms are shown as a function of the 
	seven variables used for the exclusivity selection described in the text, 
	plus the missing $P_x$ and $P_y$, in order left to right and top to bottom. 
	}
\label{fig:CohExcCuts}
\end{figure}

\begin{figure}[tbp!]
\center
\includegraphics[trim=20 10 20 5,clip,width=15cm]{fig3/all_incoh_exc_cuts.pdf}
\caption{Distributions of the incoherent photon production events before 
	(black) and after (blue line filled
	in Grey) the exclusivity cuts used to select coherent DVCS represented by the 
	red dashed lines. The histograms are shown as a function of the 
	seven variables used for the exclusivity selection described in the text, 
	plus the missing $P_x$ and $P_y$, in order left to right and top to bottom.}
\label{fig:IncExcCuts}
\end{figure}
